Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-09-11 main.tex
EIF section made better, added a reference for it as well Started on private filter description

The localisation scenario we consider in this work is that of model-based self-navigation using range-only sensors. We consider localisation in the two-dimensional case for simplicty, but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & dx & y & dy
    \end{bmatrix}^\top\,.
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \mean{x}_k = f(\mean{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $w_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    z_k = h_i(\mean{x}_k)+\rv{v}_k\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_k \sim \mathcal{N}(\vec{0},\mat{R})$, while the measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        x_{i} & y_{i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\mean{x}_k) &= \left\|
        \begin{bmatrix}
            \mean{x}_{k[0]} \\
            \mean{x}_{k[2]}
        \end{bmatrix}
        - \vec{s}_{i}\right\| \\
        &= \sqrt{(\mean{x}_{k[0]}-x_{i})^2 + (\mean{x}_{k[2]}-y_{i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn the estimated state $\mean{x}_k$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$ or their measurements $z_k$ at any time $k$.

@@

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within range of a navigator broadcast. We loosely propose that this scenario can be handled by pre-defined sensor subsets during an initial step by a trusted party. This has been captured in Figure \ref{fig:sensor_subsets}.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is recieved equal by all sensors. This means the navigator may not send different information to individual sensors during a single timestep.

@@

\subsection{Extended Information Filter}
EIF is a reformulation of the EKF where the update step is reduced to sums. Makes it especially easier for processing multiple measurements.

Prediction step can be computed in either the EIF form or the equivalent EKF state and covariance can be computed before performing the prediction step in EKF form and converting the result back.

% The Extended Information Filter (EIF) is an algebraic reformulation of the Extended Kalman Filter (EKF) such that the update step is reduced to a simple summation. Given process and measurement models \eqref{} and \eqref{}, the EIF update equation is given by
% \begin{equation}
%     a
% \end{equation}

@@

\section{Private Localisation with Privacy-Preserving Sensors} \label{sec:private_localisation}
% Explain it in as an overview. How is the aggregation scheme used, what does this require from the measurement model, why can this be a problem for normal distance sensors?

@@

Using the EIF and distance sensors, we see that the required computation of measurement vector and matrices is reduced to computing $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$, for one-dimensional $R$ and jacobian $H$, at each sensor. As this requires current state information $x$, relevant attributes need to be sent to the sensors.

>>>>

The localisation scenario we consider in this work is that of model-based self-navigation using range-only sensors. We consider localisation in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & dx & y & dy
    \end{bmatrix}^\top\,.
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \mean{x}_k = \vec{f}(\mean{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $w_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    z_k = h_i(\mean{x}_k)+\rv{v}_k\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_k \sim \mathcal{N}(\vec{0},\mat{R})$, while the measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        x_{i} & y_{i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\mean{x}_k) &= \left\|
        \begin{bmatrix}
            \mean{x}_{k[0]} \\
            \mean{x}_{k[2]}
        \end{bmatrix}
        - \vec{s}_{i}\right\| \\
        &= \sqrt{(\mean{x}_{k[0]}-x_{i})^2 + (\mean{x}_{k[2]}-y_{i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn the estimated state $\mean{x}_k$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$ or their measurements $z_k$ at any time $k$.

@@

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within the range of a navigator broadcast. We loosely propose that this scenario can be handled by pre-defined sensor subsets during an initial step by a trusted party. This has been captured in Figure \ref{fig:sensor_subsets}.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is received equal by all sensors. This means the navigator may not send different information to individual sensors during a single timestep.

@@

\subsection{Extended Information Filter}
The Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982} is a reformulation of the Extended Kalman Filter (EKF), such that the update equations are simplified to sums of measurement information from each sensor. This requires the conversion of state estimate $\mean{x}_k$ and estimate covariance $\mat{P}_k$ to the information vector and matrix
\begin{equation}
    \mean{y}_k = \mat{P}_k^{-1}\mean{x}_k \quad\textrm{and}\quad \mat{Y}_k = \mat{P}_k^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ measurements at time $k$, given process and measurement models \eqref{eqn:process_model} and \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \mean{y}_k = \mean{y}_{k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} \mat{R}^{-1}_i (z_{k,i} - h_i(\mean{x}_k) + \mat{H}_{k,i}\mean{x}_k) \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_k = \mat{Y}_{k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} \mat{R}^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobian
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\mean{x}_k} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $1 \leq i \leq n$. The filter's prediction step can be computed by converting information vector $\mean{y}_k$ and matrix $\mat{Y}_k$ back to state estimate and covariance 
\begin{equation}
    \mean{x}_k = \mat{Y}_k^{-1}\mean{y}_k \quad\textrm{and}\quad \mat{P}_k = \mat{Y}_k^{-1}\,,
\end{equation}
before using the normal EKF prediction equations
\begin{equation}
    \mean{x}_k = \vec{f}(\mean{x}_{k-1})
\end{equation}
and
\begin{equation}
    \mat{P}_k = \mat{F}_k\mat{P}_{k-1}\mat{F}_k^\top\,,
\end{equation}
with Jacobian
\begin{equation}
    \mat{F}_{k} = \left.\frac{\partial \vec{f}}{\partial \vec{x}}\right|_{\mean{x}_k}\,.
\end{equation}

@@

\section{Private Localisation with Privacy-Preserving Sensors} \label{sec:private_localisation}
Given our definitions for the aggregation scheme in Section \ref{sec:our_scheme} and preliminaries in Section \ref{sec:loc_prelim}, we can now construct a localisation method meeting our problem formulation in Section \ref{sec:problem_statement}. We consider a navigator in the presence of $n$ range sensors, running a local filter with update steps computed by equations \eqref{eqn:eif_info_vec_update} and \eqref{eqn:eif_info_mat_update}. Received information from each sensor $i$ consists of the measurement vector
\begin{equation}
    \vec{i}_i = \mat{H}^\top_{k,i} \mat{R}^{-1}_i (z_{k,i} - h_i(\mean{x}_k) + \mat{H}_{k,i}\mean{x}_k)
\end{equation}
and the measurement matrix
\begin{equation}
    \mat{I}_i = \mat{H}^\top_{k,i} \mat{R}^{-1}_i \mat{H}_{k,i}\,,
\end{equation}
at each time step $k$. All sensitive sensor information is captured exclusively in the filter by the Products $\vec{i}_i$ and $\mat{I}_i$, namely their measurements $z_{k,i}$, measurement variances $\mat{R}_i$ and locations, captured in measurement functions $h_i$ and Jacobians $\mat{H}_{k,i}$. Encryption of $\vec{i}_i$ and $\mat{I}_i$ at the sensors would, however, disclose the location information of $\mean{x}_k$, required for computing Jacobians $\mat{H}_{k,i}$ in \eqref{eqn:measurement_jacobian}.

% Explain it in as an overview. How is the aggregation scheme used, what does this require from the measurement model, why can this be a problem for normal distance sensors?

@@
